---
title: "ASN5.2. Modeling for Inference"
author: "2023 FIRE298 Sustainability Analytics"
format: html
editor: visual
---

In this assignment, you will use the difference-in-difference model framework to estimate the impact of the treatment variable.

### **Part 1. Explaining the Data**

Instructions: Run the script below to download and declare the required packages.

```{r}
#| message: false
#install.packages("tidyverse")
library("tidyverse")
```

Instructions: Upload your team's final dataframe into Posit Cloud. Use the read.csv function to upload the dataframe into the Posit Cloud Environment.

```{r}
read_csv("loadsdatatable.csv") -> df
```

**Question 1: Which variable indicates the treatment *group*? If your dataset does not have one yet, please create it using the mutate and ifelse function (2 points)**

Answer: The treatment group is whether a mine is a cobalt mine or not, which is represented by the column re_Co (binary variable that stores whether a mine produces cobalt or not)

**Question 2: Which variable indicates the treatment *period?* If your dataset does not have one yet, please create it using the mutate and ifelse function (2 points)**

Answer: The treatment period is from 1997 to 2022 and is represented by the column active (binary variable which stores whether a mine was active or not for each year between 1997 and 2022).

**Question 3: Run a DiD regression with the treatment group** $d_i$**, treatment period** $t$**, and its interaction** $d_{it} \times t$ **to estimate their impact on the outcome variable. This regression can be represent by the equation** $Y_{it}=\beta_0+\beta_1 d_i+\beta_2 t + \gamma (d_{it}\times t) + \epsilon_{it}$

**Afterward, print the regression results by using the function summary(*ModelName*). An example of this can be found in the lecture slides 2023 FIRE298 Week 12 Part 2 Slide 5.**

```{r}
summary(c1<- lm(n ~ active + re_Co + active:re_Co, data=df))

```

**Based on the regression, what is the value of** $\beta_0$ **? (1 point)**

Answer: 0.309

**Based on the regression, what is the value of** $\beta_1$ **? (1 point)**

Answer: -0.217

**Based on the regression, what is the impact of the treatment on the treatment group, relative to the control group, represented by** $\gamma$ **? (1 point)**

Answer: The treatment effect (represented by $\gamma$) is 0.231, which indicates that for every year a cobalt mine is open, it is predicted to have 0.231 more conflicts on average when compared to other types of mines

**Based on the regression, is the treatment effect statistically different from 0? (1 point)**

Answer: The $\gamma$ variable has three asterisks in the regression, which means it is statistically significant for even extremely low p-values.

### **Part 2. DiD with Fixed Effects**

**Question 4: Run a DiD model with the treatment group** $d_i$**, treatment period** $t$**, its interaction** $d_{it} \times t$, **and all other** **numerical variables in your data to estimate the treatment effect. What does the regression table look like? (1 point)**

```{r}
summary(c1<- lm(n ~ active + re_Co + active:re_Co + year_do + year_de + year + active_97_22, data=df))

```

**Question 5: Based on the regression in Question 4, what is the** **treatment effect? Is the treatment effect statistically different from 0? (2 points)**

The treatment effect (represented by $\gamma$ ) is 0.164, which indicates that cobalt mines are predicted to have 0.164 more conflicts on average when compared to other types of mines for every year they are open. The $\gamma$ variable has three asterisks in the regression, so it is statistically significant, even at extremely low p-values.

```         
```

**Question 6: Run a DiD model the treatment group** $d_i$**, treatment period** $t$**, its interaction** $d_{it} \times t$, **and all other numerical, binary, and categorical variables. What does the regression table look like? (2 points)**

Instructions: A binary variable takes the form of 0 or 1. A categorical variable is a character with only a few unique values.

The efficient way to include categorical variables in a regression is to use the felm( ) function from the lfe package instead of the lm( ) function. The function has the format:

``` r
ModelName<- felm(OutcomeVariable ~ NumericalVariables + BinaryVariables | Categorical Variables, DataFrame)
```

The summary(ModelName) function is then used to print out the regression model.

```{r}
#| message: false
#install.packages("lfe")
library("lfe")
```

Write your model below and run it.

```{r}
modelName <- felm(n ~ active + re_Co + active:re_Co + year_do + year_de + year + active_97_22 + re_Sn + re_Ta | parties + type, data = df)
summary(modelName)
```

**Question 7: Based on the regression in Question 6, what is the** **treatment effect? Is the treatment effect statistically different from 0? (2 points)**

The treatment effect (represented by $\gamma$) is 0.148, which indicates that cobalt mines are predicted to have 0.148 more conflicts on average when compared to other types of mines for every year they are open. The $\gamma$ variable has three asterisks in the regression, so it is statistically significant, even at extremely low p-values.

**Question 8: Create more binary or categorical variables by multiplying binary variables or combining categorical variables together as you see fit. Explain what real-world factor each new variable represents. (2 points)**

Instructions: You created new variables by multiplying binary variables and combining categorical variables in asn-6-4-key.qmd Part 2. You explained what each variable represents in asn-6-4-key.qmd Part 1 Questions 4, 5, and 6.

```{r}
df2 <- df %>%
  mutate(re_TaSn = re_Ta * re_Sn) %>%
  mutate(party_and_type = str_c(parties, type, sep = " - "))
  
```

Answer:

re_TaSn - this new column represents the mines that contain both Tantalum and Tin.

party_and_type - this new column combines the parties and type categorical variables into one column

**Question 9: Run a DiD model with the treatment group** $d_i$**, treatment period** $t$**, its interaction** $d_{it} \times t$, **and all other numerical, binary, and categorical variables that you have created in Question 8. What does the regression table look like? (2 points)**

```{r}
modelName1 <- felm(n ~ active + re_Co + active:re_Co + year_do + year_de + year + active_97_22 + re_TaSn | party_and_type, data = df2)
summary(modelName1)
```

**Question 10: Based on the regression in Question 9, what is the** **treatment effect? Is the treatment effect statistically different from 0? (2 points)**

Answer: The treatment effect (represented by $\gamma$) is 0.173, which means that cobalt mines are predicted to have 0.173 more conflicts on average when compared to other types of mines for every year they are open. The $\gamma$ variable has three asterisks in the regression, so it is statistically significant, even at extremely low p-values.

**Question 11: Based on your intuition**, **which DiD regression was the best estimator of the treatment effect?** **Please explain why. (2 points)**

Answer: Based on our intuition, the DiD regression completed under question 6 is likely the best estimator of the treatment effect. This is because it controls for consequential numerical and categorical variables (such as active_97_22 and parties), but does not go overboard and include any unnecessary variables like the DiD regression completed under question 9.

### **Part 3 Thinking about causal inference**

**Question 12: A credible DiD analysis assumes that no other factor impacts the changes in the treatment group that have not been controlled. Based on your opinion, what factors can change the treated outcome that has not been accounted for? (10 points)**

Answer: Factors like annual precipitation, population, and income levels/access to basic services in the buffer zone of the mine may all influence the amount of conflict in that zone.

**Question 13: Had you had more time, what additional variables (that you know exist) could you add to your data to take into account external factors that influence the treatment outcome? Please provide a link to each of the data sources. (4 points)**

Answer: If we had more time, we would have likely included the control variable of annual precipitation, which can be found here: <https://search.earthdata.nasa.gov/search?q=IMERG>

We would have also controlled for population, which can be found here: <https://data.humdata.org/dataset/cod-ps-cod?>

Lastly, we would have likely also controlled for poverty rates, which can be found here: [https://www.oxfam.org/en/what-wedo/countries/democratic-republic-congo](https://www.oxfam.org/en/what-we-do/countries/democratic-republic-congo)

You have reached the end of this assignment. Commit and push your updates back into the GitHub repository. You can find the step-by-step tutorial [here.](https://drive.google.com/drive/u/0/folders/1eSMfNoV_ilp4OnID2Ql3YT2qXkhCS7Vm)
